---
title: "Peer Assessment 1"
author: "cafkhami"
date: "Thursday, March 12, 2015"
output: html_document
---

This is my R markdown file for the first Peer Assessment in the Reproducable Research class

First we will set the WD and read in the data and set all NA values to 0
```{r, echo= TRUE}
setwd("~/coursera/RR/RepData_PeerAssessment1")
data <- read.csv("activity.csv")
check <- is.na(data$steps)
data.NA <- data
data.NA[check,1] <- 0
```

Now we will calculate the total number of steps in each day
```{r, echo=TRUE}
DaySum <- tapply(data.NA$steps, data.NA$date, sum)
```

Let's view the data in a histogram
```{r, echo=TRUE}
hist(DaySum, col = "red", xlab = "Steps Per Day")
```

The median and mean steps taken per day:
```{r, echo=TRUE}
mean(DaySum)
median(DaySum)
```

Now lets look at the average activity for a given interval.  First we will calculate the average and then plot it.
```{r, echo=TRUE}
IntAvg <- tapply(data.NA$steps, data.NA$interval, mean)
plot(names(IntAvg), IntAvg, type ="l", xlab = "Interval", ylab = "Average Steps", main = "Average Steps Throughout the Day")
```

The total number of missing values in the dataset
```{r, echo=TRUE}
sum(check)
```
let's fill in those missing data points in a new dataset called `data.comp` with the average of that interval for the rest of the dataset
```{r, echo=TRUE}
data.comp <- data
data.comp[check,1] <- IntAvg[as.character(data[check,3])]
```
And lets view the histogram, mean and median of this data:
```{r, echo=TRUE}
DaySum.comp <- tapply(data.comp$steps, data.comp$date, sum)
hist(DaySum.comp, col = "red", xlab = "Steps Per Day")
mean(DaySum.comp)
median(DaySum.comp)
```

As you can see there is a significant impact on the data when estimated values are used instead of missing values:

- The Mean and Median both increase
- the number of 0 step days decreases
- the number of steps per day increases accross the board.

Now we will look at the differences between activity on weekdays vs weekends
